---
title: "Bike Sharing Project"
author: "Jacob Paul"
date: "2024-10-12"
output: html_document
---

## 📌 Project Goal

This report replicates the steps in Question 13-C using R Markdown for reproducibility. It includes data loading, transformations, linear regression, and summary interpretation on the bike-sharing dataset.

---

## 🔹 Load the Data

```{r}
day <- read.csv("data/day.csv")
head(day)
```

---

## 🔹 Convert Numerical to Categorical Labels

### Convert `season` from numeric to descriptive labels

```{r}
day$season <- as.character(day$season)
day$season[day$season == "1"] <- "springer"
day$season[day$season == "2"] <- "summer"
day$season[day$season == "3"] <- "fall"
day$season[day$season == "4"] <- "winter"
print(unique(day$season))
```

### Convert `weathersit` from numeric to descriptive labels

```{r}
day$weathersit <- as.character(day$weathersit)
day$weathersit[day$weathersit == "1"] <- "Good"
day$weathersit[day$weathersit == "2"] <- "Mist"
day$weathersit[day$weathersit == "3"] <- "Bad"
day$weathersit[day$weathersit == "4"] <- "Severe"
print(unique(day$weathersit))
```

---

## 🔹 Convert Categorical Variables to Factors

```{r}
day$season <- as.factor(day$season)
day$holiday <- as.factor(day$holiday)
day$workingday <- as.factor(day$workingday)
day$weathersit <- as.factor(day$weathersit)
str(day)
```

---

## 🔹 Build Linear Model (Q13-C)

```{r}
model_3 <- lm(cnt ~ season + holiday + workingday + weathersit + atemp + hum + windspeed + registered, data = day)
summary(model_3)
```

---

## 🔹 Interpretation

### Model Formula (Key Predictors Only):

```
cnt = 575.19 + 1378.93 × atemp - 363.99 × hum - 637.75 × windspeed + 1.16 × registered
```

These coefficients show that:
- `atemp` and `registered` have strong positive relationships with bike count
- `hum` and `windspeed` negatively affect count

---

## 📈 Conclusion

This model explains 96.5% of the variance (Adjusted R² = 0.9653), making it highly predictive for bike rental demand based on environmental and user-based features.
